
# Exercise 2 -------------------------------------------------------------------
Control flow operators, Functions, Packages
Please try to complete tasks listed below. Type your code in each section.



clean up global environment
```{r}
rm(list = ls())
```


### Task 1 ---------------------------------------------------------------------

Air temperature above 30 degrees might be considered as "too hot". 
Temperature below 5 degrees is "too cold". Anything in between is "good".
Use conditions to take a value of air temperature "Temp" and print
the corresponding condition: "good" or "too hot" or "too cold".

```{r}
Temp <- 33    # Try different values to test your code
if (Temp > 30) {
  cat("The temperature is too hot.")
} else if (Temp < 5) {
  cat("The temperature is too cold.")
} else {
  cat("The temperature is good.")
}
```



### Task 2 ---------------------------------------------------------------------

Use the code from Task 1 and convert it into a custom function that takes 
as an input air temperature "Temp" and outputs the corresponding condition.
Then use a loop with a function to get conditions for the following values
of the air temperature and store them as a vector "conditions"

```{r}
# Define custom function to get condition based on air temperature
get_condition <- function(Temp) {
  if (Temp > 30) {
    return("too hot")
  } else if (Temp < 5) {
    return("too cold")
  } else {
    return("good")
  }
}

# Define vector of air temperatures
temps <- c(0, 15, 35, 20, -10, 44)

# Use loop with function to get conditions for each temperature
conditions <- c()
for (temp in temps) {
  condition <- get_condition(temp)
  conditions <- c(conditions, condition)
}

# Print the conditions vector
conditions
```




### Task 3 ---------------------------------------------------------------------

Make a code to re-create a sequence of first 30 Fibonacci numbers
Every next Fibonacci number equals to the sum of two previous numbers
1, 1, 2, 3, 5, 8, ...

```{r}
# Define the first two Fibonacci numbers
fib <- c(1, 1)

# Loop to generate the next 28 Fibonacci numbers
for (i in 3:30) {
  next_fib <- fib[i-1] + fib[i-2]
  fib <- c(fib, next_fib)
}

# Print the sequence of Fibonacci numbers
fib
```



### Task 4 ---------------------------------------------------------------------

Do a Google search and find ready-made package and function to output
Fibonacci numbers. Get first 30 numbers and compare them to your own result.

# Install the numbers package if not already installed
install.packages("numbers")


```{r}
# Load the numbers package
library(numbers)

# Generate the first 30 Fibonacci numbers
fib <- numeric(30)
for (i in 1:30) {
  fib[i] <- fibonacci(i)
}

expected_result <- c(1, 1, 2, 3, 5, 8, 13, 21, 34, 55, 89, 144, 233, 377, 
                     610, 987, 1597, 2584, 4181, 6765, 10946, 17711, 28657,
                     46368, 75025, 121393, 196418, 317811, 514229, 832040)
# Compare the generated Fibonacci numbers with the expected result
identical(fib, expected_result)
```



### Task 5 ---------------------------------------------------------------------

Take a sequence of normally distributed random numbers as below and 
count how many times the value in sequence is greater than 2 or less than -2.
Make 4 versions of the code by using: 
(1) for loop; (2) while loop; (3) repeat loop; (4) no loop

X <- rnorm(1000)

## 1 for loop

```{r}
X <- rnorm(1000)
count <- 0

for (i in 1:length(X)) {
  if (X[i] > 2 || X[i] < -2) {
    count <- count + 1
  }
}

count
```


## 2 while loop

```{r}
X <- rnorm(1000)
count <- 0
i <- 1

while (i <= length(X)) {
  if (X[i] > 2 || X[i] < -2) {
    count <- count + 1
  }
  i <- i + 1
}

count
```

## 3 repeat loop
```{r}
X <- rnorm(1000)
count <- 0
i <- 1

repeat {
  if (X[i] > 2 || X[i] < -2) {
    count <- count + 1
  }
  i <- i + 1
  
  if (i > length(X)) {
    break
  }
}

count
```

## 4 no loop 
```{r}
X <- rnorm(1000)
count <- sum(X > 2 | X < -2)

count
```
