
# Exercises for week 3 ---------------------------------------------------------
Data loading, data visualisation, descriptive statistics


Please try to complete tasks listed below. Type your code and results 
interpretations in each section.



clean up global environment
```{r}
rm(list = ls())
```


### Task 1 ---------------------------------------------------------------------

Check your working directory (and/or adjust your working directory) to ensure
that data file "week03_data_oxygen.csv" is placed in the working directory.
Then load the data contained in the "week03_data_oxygen.csv" file.
Provide comments about the data set.

Data information: The oxygen distribution gives important information 
on the general biochemistry of marine life. The data measures the oxygen 
saturation (mg/L^10) at 22 marine parks in Australia. 

```{r}
# load data
oxygen <- read.csv("data_oxygen.csv")
# check data
head(oxygen)
str(oxygen)
summary(oxygen)
dim(oxygen)
```



### Task 2 ---------------------------------------------------------------------
Visualise data distribution by histogram and boxplot
Take care about a clear and informative titles and axis labels for all 
your visualisations.
Provide your comments about the shape of the distribution as a comment.

```{r}
# histogram
hist(oxygen$Oxygen, main = "Oxygen distribution", xlab = "Oxygen saturation (mg/L^10)", ylab = "Frequency")
```

```{r}
# histogram
# breaks, bins
hist(oxygen$Oxygen, breaks= 10, main = "Oxygen distribution", xlab = "Oxygen saturation (mg/L^10)", ylab = "Frequency")
```

```{r}
# histogram
# breaks, bins
hist(oxygen$Oxygen, breaks= seq(6,15,2), main = "Oxygen distribution", xlab = "Oxygen saturation (mg/L^10)", ylab = "Frequency")
```



```{r}
# histogram
hist(log(oxygen$Oxygen), breaks= 6, main = "Oxygen distribution", xlab = "Oxygen saturation (mg/L^10)", ylab = "Frequency")
```

```{r}
# boxplot
boxplot(oxygen$Oxygen, main = "Oxygen distribution", ylab = "Oxygen saturation (mg/L^10)")
```



### Task 3 ---------------------------------------------------------------------

Calculate descriptive statistics for the data. 
Get both options for centrality and dispersion measures. 
Comment what measures are more appropriate for these data, provide 
an interpretation for these measures.


```{r}
# mean
mean(oxygen$Oxygen)
# median
median(oxygen$Oxygen)
# range
range(oxygen$Oxygen)
# variance
var(oxygen$Oxygen)
# standard deviation
sd(oxygen$Oxygen)
# interquartile range
IQR(oxygen$Oxygen)
```

```{r}
summary(oxygen$Oxygen)
```

```{r}
library(moments)
moments::skewness(oxygen$Oxygen)
moments::kurtosis(oxygen$Oxygen)
```




```{r}
```