



# Exercise 1 -------------------------------------------------------------------
Data types and data structures, indexing


Please try to complete tasks listed below. Type your code in each section.


# clean up global environment
rm(list = ls())



### Task 1 ---------------------------------------------------------------------

Create a character vector. 
Use three different methods to get the character vector of length 4.


```{r}
# Method 1: Using the c() function
my_vector <- c("a", "b", "c", "d")
my_vector
```

```{r}
# Method 2: Using the paste() function
my_vector <- paste("a", "b", "c", "d", sep = ",")
my_vector
```

```{r}
# Method 3: Using the strsplit() function
my_string <- "abcd"
my_vector <- strsplit(my_string, "")[[1]]
my_vector
```




### Task 2 ---------------------------------------------------------------------

Create a sequence of 8 numbers, then apply a mathematical
operation that should result in the following output:
[1]    8   16   32   64  128  256  512 1024
```{r}
# Create a sequence of 8 numbers
x <- seq(1, 8)

# Apply a mathematical operation
y <- 2^(x+2)

# Print the output
print(y)
```






### Task 3 ---------------------------------------------------------------------

Create a data frame as below 
   x y z
 1 a 1 5
 2 b 2 6
 3 c 3 7

```{r}
# Create the data frame
df <- data.frame(x = c("a", "b", "c"),
                 y = c(1, 2, 3),
                 z = c(5, 6, 7))

# Print the data frame
df
```


Use different methods of indexing to extract values "a" and "c" together.
You should get at least 4 different options.

```{r}
# Extract "a" and "c" using row indexing with logical conditions
subset_df <- df[df$x %in% c("a", "c"), ]
desired_values <- subset_df$x
desired_values
```

```{r}
# Extract "a" and "c" using row and column indexing with numeric indices
desired_values <- df[c(1, 3), "x"]
desired_values
```

```{r}
# Extract "a" and "c" using row and column indexing with logical conditions
desired_values <- df[df$x == "a" | df$x == "c", "x"]
desired_values
```

```{r}
# Extract "a" and "c" using the subset() function
subset_df <- subset(df, x %in% c("a", "c"))
desired_values <- subset_df$x
desired_values
```



### Task 4 ---------------------------------------------------------------------


Run the code below to load data set about passengers of Titanic.

```{r}
data(Titanic)
Titanic
```

Check the help file with data description
```{r}
help(Titanic)
```

Use indexing to extract the number of crew member of both genders 
survived and died - to get the total count of crew members on Titanic
```{r}
# Load the Titanic data set
data(Titanic)
# change to data frame
Titanic <- as.data.frame(Titanic)
Titanic
```

```{r}
# Extract the number of crew members who survived and died
crew_survived <- sum(Titanic[Titanic$Class == "Crew" & Titanic$Survived == "Yes", "Freq"])
crew_died <- sum(Titanic[Titanic$Class == "Crew" & Titanic$Survived == "No", "Freq"])

# Calculate the total count of crew members
crew_total <- sum(Titanic[Titanic$Class == "Crew", "Freq"])

# Print the results
cat("Number of crew members who survived:", crew_survived, "\n")
cat("Number of crew members who died:", crew_died, "\n")
cat("Total count of crew members:", crew_total, "\n")
```


### Task 5 ---------------------------------------------------------------------



Run the code below to load data set about car performance.
```{r}
data(mtcars)
mtcars
```

Check the help file with data description

help(mtcars)

Variable "vs" do not look as truly numerical, it has a meaning
(V-shape or Straight) but not a numerical value. Hence, it should be 
converted to factor. Create a new variable in the data set "vs.f" 
with corresponding factor values - V-shape or Straight.

Hint: check examples for function "factor()"
help(factor)

```{r}
# Convert "vs" variable to factor
mtcars$vs.f <- factor(mtcars$vs, levels = c(0, 1), labels = c("Straight", "V-shape"))

# Print the updated data set
mtcars
```




```{r}
```



